Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-10-07 GWLDCII.tex
few small changes in read through of sec 5.

The results of Section~\ref{sec GN algorithm} allow us to relate the configuration of propagators in a Wilson loop diagram $W$ to minors of $C(W)$, $R(W)$.  In this section fix an arbitrary ordering of the propagators, equivalently the rows of $C(W)$.\note{There's some grammatical issues in this paragraph and I'm not sure what it's supposed to be saying.}

@@

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP,[n])$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in equation \eqref{C(W) dfn} with a fixed ordering on $\cP$ (see Section~\ref{section WLD background})\note{This was written before the exposition above; now I think we can simplify it to "Let $W$ be a WLD and $C(W)$ its associated matrix?}.
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{0}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{0} = I_n$.)
	  %\item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be the determinant of the $|S_i|\times |S_i|$ minor of $C(W)$ with rows indexed by $S_i$ and columns indexed by $I_i(S_i)$.  If $S_i = \emptyset$, set $r_i = 1$. %$\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Viewing the determinant as a sum over permutations, and combining this with Corollary~\ref{lem basis as perm}, we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, namely \hlfix{$I_i(p)$}{do we mean $I_i(p)$ or $V(p)$ here? if the first, then the last part of the sentence is wrong}, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

Let $S = \cP_{in}(p)$ be the set of propagators inside $p$ in the $<_i$ order.
By Remark \ref{rmk algorithm locally same}, $I_i$ and $I_{i+1}$ can only differ once $p$ contributes to $I_{i+1}$, so $I_i(q) = I_{i+1}(q)$ for each $q \in S \backslash \{p\}$. Thus if a propagator contributes $m$ in $I_i$ then it must lie outside $p$.

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} we have $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$. However, all propagators in $S$ contribute to $I_{i}$ strictly before $m$ by \st{the previous paragraph}{\color{violet}Remark {clockwise ordering rem}}\note{The purple edit isn't true: a priori a propagator inside $p$ could lie on edge $m$ and contribute $m$. The previous paragraph explains why this doesn't happen.} so, after permuting rows and columns as needed, the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A${\color{violet}. If we remove the propagators that contribute variables to the minor $C$,}\st{ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$,} we get a diagram $W'$ with fewer propagators for which $\det A | \Delta_{I^{W'}_i}$ still holds, hence $f|\Delta_{I^{W'}_i}$.  This contradicts the minimality of our choices. \sanote{trying to make clear that the minor changes with the diagram. May have just made it too notation heavy.}

@@

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ contribute to $I_i$ strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form {\color{violet}(when reordered according to how they contribute to $I_i$)}
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $<_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}=\det A\det B$.  This yields a contradiction as in Case 1: $f|\det A$ and removing a propagator not in $S$ gives a smaller diagram with a factor of degree $\geq 3$, which contradicts our minimality assumption.\note{I rephrased this last sentence quite a bit because earlier things had gotten rephrased} \sanote{reads fine to me.}

@@

\item \sanote{changed $b$ to $a+1$ here.} Suppose propagators $p$ and $q$ are supported on a common edge  $a$, with $I_i(p)=a$ and $I_i(q)=a+1$.  Let $x_{p,a},x_{p,a+1},x_{q,a},x_{q,a+1}$ be the associated variables in $C(W)$. For any fixed bijection $\sigma$ from $\cP\setminus\{p,q\}$ to $I_i \setminus\{a,a+1\}$ for which each propagator is supported on its image under the bijection, we can extend $\sigma$ to a bijection from propagators to $I_i$ in two ways: either $p\mapsto a$ and $q\mapsto a+1$ or $p\mapsto a+1$ and $q\mapsto a$.  The sum of the contributions of all these bijections to $\Delta_{I_i}$ is therefore the product of $x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a}$ {\color{violet} By the arguments in the proof of the claim in part 2 of this theorem} \note{Could this use a ref back to where we discuss this kind of arugment in more detail.}\sanote{see violet?} with the minor coming from $\cP\setminus\{p, q\}$ and $I_i \setminus \{a,a+1\}$.  Since there is no cancellation of terms in the expansion of $\Delta_{I_i}$, if any other terms appear then they will cause a factor which is not in the form described in the previous part.  Therefore no such terms exist and $x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a}$ is a factor of $\Delta_{I_i}$.

@@

\item If $S_i = \emptyset$ then $r_i = 1$ and we immediately have $r_i | \Delta_{I_i}$; otherwise, let $p\in S_i$. Note that $\Delta_{I_i}$ is homogeneous linear in the variables of the row corresponding to $p$.  By part (2), either exactly one variable in the row corresponding to $p$ appears in $\Delta_{I_i}$ and this variable is a factor of $\Delta_{I_i}$, or exactly two variables from the row corresponding to $p$ appear in $\Delta_{I_i}$ and they appear as part of a quadratic factor.  In the first case let the variable be $x$. Then $x$ is a factor of $\Delta_{I_i}$. {\color{violet} Therefore, $x$ must correspond to entry in the row of $p$ and column of $I_i(p)$, i.e. $x = x_{p, I_i(p)}$.} \st{and so in particular the monomial in $\Delta_{I_i}$ corresponding to the bijection between propagators and vertices of $I_i$ associates the column of $x$ to $p$. Thus $x$ also appears in $r_i$ and s} Since the matrix for $r_i$ is a minor of the matrix for $\Delta_{I_i}$ \st{and every term in $\Delta_{I_i}$ involves $x$,} we also have that \st{every term in $r_i$ involves $x$ so}{\color{violet}$x$ is a factor of $r_i$.} \st{$x$ is a factor of both $r_i$ and $\Delta_{I_i}$ and is the only variable from this row in either polynomial.}

>>>>

The results of Section~\ref{sec GN algorithm} allow us to relate the configuration of propagators in a Wilson loop diagram $W$ to minors of $C(W)$ contributing to $R(W)$.  In this section, we fix an arbitrary ordering on the propagators of W, i.e we impose and ordering on the rows of $C(W)$.\note{There's some grammatical issues in this paragraph and I'm not sure what it's supposed to be saying.} \sanote{tweaked? better?}

@@

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP, [n])$ be a Wilson loop diagram and $C(W)$ be its associated matrix with a fixed ordering imposed on the rows.%Let $W = (\cP,[n])$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in equation \eqref{C(W) dfn} with a fixed ordering on $\cP$ (see Section~\ref{section WLD background})
\note{This was written before the exposition above; now I think we can simplify it to "Let $W$ be a WLD and $C(W)$ its associated matrix?}\sanote{agreed. changed. previous text as a comment}
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{0}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{0} = I_n$.)
	  %\item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be the determinant of the $|S_i|\times |S_i|$ minor of $C(W)$ with rows indexed by $S_i$ and columns indexed by $I_i(S_i)$.  If $S_i = \emptyset$, set $r_i = 1$. %$\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

@@

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Viewing the determinant as a sum over permutations, and combining this with Corollary~\ref{lem basis as perm}, we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, namely \hlfix{$I_i(p)$}{do we mean $I_i(p)$ or $V(p)$ here? if the first, then the last part of the sentence is wrong}\sanote{I read this as saying $p \rightarrow I_i(p) \in V(p) \cap I_i$}, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

Let $S = \cP_{in}(p)$ be the set of propagators inside $p$ in the $<_i$ order.
By Remark \ref{rmk algorithm locally same}, $I_i$ and $I_{i+1}$ can only differ once $p$ contributes to $I_{i+1}$, so $I_i(q) = I_{i+1}(q)$ for each $q \in S \backslash \{p\}$. Thus if a propagator contributes $m$ in $I_i$ then it must lie outside $p$. {\color{violet} Furthermore, by Remark \ref{clockwise ordering rem} and Lemma \ref{no fourth vertex}, if $I_i(q) = m$ then all propagators in $S$ must contribute before $q$.}

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} we have $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$. However, all propagators in $S$ contribute to $I_{i}$ strictly before $m$ by \st{the previous paragraph}{\color{violet}Remark \ref{clockwise ordering rem}}\note{The purple edit isn't true: a priori a propagator inside $p$ could lie on edge $m$ and contribute $m$. The previous paragraph explains why this doesn't happen.}\sanote{it doesn't actually. It says nothing explicitly about propagator order. I'm trying (erroneously) to make this explicit.} so, after permuting rows and columns as needed, the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore {\color{violet} up to a sign, }$\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A${\color{violet}. If we remove the propagators that contribute variables to the minor $C$,}\st{ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$,} we get a diagram $W'$ with fewer propagators for which $\det A | \Delta_{I^{W'}_i}$ still holds, hence $f|\Delta_{I^{W'}_i}$.  This contradicts the minimality of our choices. \sanote{trying to make clear that the minor changes with the diagram. May have just made it too notation heavy.}

@@

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ contribute to $I_i$ strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has, {\color{violet}after a possible rearrangement of rows, } the form 
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $<_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of \hlfix{$\Delta_{I_i}=\det A\det B$}{$\det A\det C$?}.  This yields a contradiction as in Case 1: $f|\det A$ and removing a propagator not in $S$ gives a smaller diagram with a factor of degree $\geq 3$, which contradicts our minimality assumption.\note{I rephrased this last sentence quite a bit because earlier things had gotten rephrased} \sanote{reads fine to me.}

@@

\item \sanote{changed $b$ to $a+1$ here.} Suppose propagators $p$ and $q$ are supported on a common edge  $a$, with $I_i(p)=a$ and $I_i(q)=a+1$.  Let $x_{p,a},x_{p,a+1},x_{q,a},x_{q,a+1}$ be the associated variables in $C(W)$. For any fixed bijection $\sigma$ from $\cP\setminus\{p,q\}$ to $I_i \setminus\{a,a+1\}$ for which each propagator is supported on its image under the bijection, we can extend $\sigma$ to a bijection from propagators to $I_i$ in two ways: either $p\mapsto a$ and $q\mapsto a+1$ or $p\mapsto a+1$ and $q\mapsto a$.  The sum of the contributions of all these bijections to $\Delta_{I_i}$ is therefore the product of $(x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a})$ \note{Could this use a ref back to where we discuss this kind of arugment in more detail.} with the minor coming from $\cP\setminus\{p, q\}$ and $I_i \setminus \{a,a+1\}$.  Since there is no cancellation of terms in the expansion of $\Delta_{I_i}$, if any other terms appear then they will cause a factor which is not in the form described in the previous part.  Therefore no such terms exist and $x_{p,a}x_{q,a+1}-x_{p,a+1}x_{q,a}$ is a factor of $\Delta_{I_i}$.

@@

\item If \hlfix{$S_i = \emptyset$}{Should we recall the definition of $S_i$ here? Not used since 5.1} then $r_i = 1$ and we immediately have $r_i | \Delta_{I_i}$; otherwise, let $p\in S_i$. Note that $\Delta_{I_i}$ is homogeneous linear in the variables of the row corresponding to $p$.  By part (2), either exactly one variable in the row corresponding to $p$ appears in $\Delta_{I_i}$ and this variable is a factor of $\Delta_{I_i}$, or exactly two variables from the row corresponding to $p$ appear in $\Delta_{I_i}$ and they appear as part of a quadratic factor.  In the first case let the variable be $x$. Then $x$ is a factor of $\Delta_{I_i}$. {\color{violet} Therefore, $x$ must correspond to entry in the row of $p$ and column of $I_i(p)$, i.e. $x = x_{p, I_i(p)}$.} \st{and so in particular the monomial in $\Delta_{I_i}$ corresponding to the bijection between propagators and vertices of $I_i$ associates the column of $x$ to $p$. Thus $x$ also appears in $r_i$ and s} Since the matrix for $r_i$ is a minor of the matrix for $\Delta_{I_i}$ \st{and every term in $\Delta_{I_i}$ involves $x$,} we also have that \st{every term in $r_i$ involves $x$ so}{\color{violet}$x$ is a factor of $r_i$.} \st{$x$ is a factor of both $r_i$ and $\Delta_{I_i}$ and is the only variable from this row in either polynomial.}
